Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-29 numerical-analysis.tex
small text changed to conclusions

Detailed analsis of simualtion results requires both the tools to ask
sophisticated questions of the data and the ability to process vast quantities of
data at high time-cadence.  As simulations grow in size and complexity, storing data
for post-processing simply becomes intractable.  To cope with this, we have
instrumented \enzo\ with the ability to conduct analysis during the course of a
simulation.  This enables analysis with extremely high time cadence (even as often
as every subcycle of the finest refinement level), without
attempting to write an entire checkpoint output to disk.  The current mechanism
for conducting analysis in \enzo\ during the course of the simulation utilizes
the same compute resources as used by the simulation itself by transferring their usage 
from \enzo\ to the analysis routines; this is often
referred to as \textit{in situ} analysis or visualization.  Utilizing a
dynamically-scheduled second set of computing resources, often referred to as
\textit{co-scheduled} analysis or visualization, provides greater flexibility and overall
throughput at the expense of simplicity.

We expose \enzo\ 's mesh geometry and fluid quantities to the analysis platform
\texttt{yt} \citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  At compile time,
\enzo\ is (dynamically or statically) linked against the Python and NumPy
libraries necessary to create proxy objects exposing the mesh geometry,
fluid quantities and particle arrays as NumPy arrays.  This information is then
passed to a special handler inside \texttt{yt}.  \texttt{yt} interprets the
mesh and fluid information and, without saving data to disk except as
explicitly requested, constructs a native representation of the in-memory state
of the simulation that appears identical to an on-disk simulation output.
\enzo\ then executes a user-provided analysis script, which is able to access the
in-memory simulation object.  Once the analysis script has returned control to
\enzo\ , the simulation proceeds.  This process can occur at either the top of the
main ``EvolveHierarchy'' loop or at the end of a timestep at the finest level,
and the frequency with which it is called is adjustable by a run-time
parameter.  During the course of conducting analysis, the simulation is halted
until the conclusion of the analysis.

@@

Where microphysical solvers or other operator-split physics calculations can be
done in Python, \texttt{yt} can serve as a driver for these calculations.  A
major feature set that is currently being developed is to pass structured
(i.e., non-fluid) information back from \texttt{yt} into \enzo\ .  For instance,
this could be the result of semi-analytic models of the growth and evolution of
star clusters, galaxy particle feedback parameters that have been influenced by
merger-tree analysis, or even spectral energy distributions that are calculated
within \texttt{yt} and provided to \enzo\ as input into radiative transfer
calculations.  Future versions will include this, as well as the ability to
dynamically allocate compute resources to \texttt{yt} such that the simulation
may proceed asynchronously with analysis (\textit{co-scheduled} analysis).
With this functionality will also come the ability to dynamically partition
data, such that spatially-decomposed operations such as volume rendering become
feasible during the course of a simulation.

@@

One of the inherent drawbacks of a grid-based fluid method is the
inability to follow the evolution of a single parcel of fluid as it
travels through the simulation volume.  To address this, Enzo has the
capability to introduce Lagrangian ``tracer particles'' into a
calculation either at the beginning of the simulation or when
restarting the calculation.  These tracer particles are put into the
simulation in a rectangular solid volume with uniform, user-specified
spacing.  Each particle's position and velocity is updated over the
course of a single timestep $\Delta t$ as follows:

>>>>

Detailed analysis of simulation results requires both the tools to ask
sophisticated questions of the data and the ability to process vast quantities of
data at high time-cadence.  As simulations grow in size and complexity, storing data
for post-processing simply becomes intractable.  To cope with this, we have
instrumented \enzo\ with the ability to conduct analysis during the course of a
simulation.  This enables analysis with extremely high time cadence (even as often
as every subcycle of the finest refinement level), without
attempting to write an entire checkpoint output to disk.  The current mechanism
for conducting analysis in \enzo\ during the course of the simulation utilizes
the same compute resources as used by the simulation itself by transferring their usage 
from \enzo\ to the analysis routines; this is often
referred to as \textit{in situ} analysis or visualization.  Utilizing a
dynamically-scheduled second set of computing resources, often referred to as
\textit{co-scheduled} analysis or visualization, provides greater flexibility and overall
throughput at the expense of simplicity.

We expose \enzo's mesh geometry and fluid quantities to the analysis platform
\texttt{yt} \citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  At compile time,
\enzo\ is (dynamically or statically) linked against the Python and NumPy
libraries necessary to create proxy objects exposing the mesh geometry,
fluid quantities and particle arrays as NumPy arrays.  This information is then
passed to a special handler inside \texttt{yt}.  \texttt{yt} interprets the
mesh and fluid information and, without saving data to disk except as
explicitly requested, constructs a native representation of the in-memory state
of the simulation that appears identical to an on-disk simulation output.
\enzo\ then executes a user-provided analysis script, which is able to access the
in-memory simulation object.  Once the analysis script has returned control to
\enzo, the simulation proceeds.  This process can occur at either the top of the
main ``EvolveHierarchy'' loop or at the end of a timestep at the finest level,
and the frequency with which it is called is adjustable by a run-time
parameter.  During the course of conducting analysis, the simulation is halted
until the conclusion of the analysis.

@@

Where microphysical solvers or other operator-split physics calculations can be
done in Python, \texttt{yt} can serve as a driver for these calculations.  A
major feature set that is currently being developed is to pass structured
(i.e., non-fluid) information back from \texttt{yt} into \enzo.  For instance,
this could be the result of semi-analytic models of the growth and evolution of
star clusters, galaxy particle feedback parameters that have been influenced by
merger-tree analysis, or even spectral energy distributions that are calculated
within \texttt{yt} and provided to \enzo\ as input into radiative transfer
calculations.  Future versions will include this, as well as the ability to
dynamically allocate compute resources to \texttt{yt} such that the simulation
may proceed asynchronously with analysis (\textit{co-scheduled} analysis).
With this functionality will also come the ability to dynamically partition
data, such that spatially-decomposed operations such as volume rendering become
feasible during the course of a simulation.

@@

One of the inherent drawbacks of a grid-based fluid method is the
inability to follow the evolution of a single parcel of fluid as it
travels through the simulation volume.  To address this, \enzo\ has the
capability to introduce Lagrangian ``tracer particles'' into a
calculation either at the beginning of the simulation or when
restarting the calculation.  These tracer particles are put into the
simulation in a rectangular solid volume with uniform, user-specified
spacing.  Each particle's position and velocity is updated over the
course of a single timestep $\Delta t$ as follows:
